---
title: "R for Data Science"
author: "Kristina Kwock"
date: '`r Sys.Date()`'
output: html_document
---

# R for Data Science

Task is to analyze the gapminder_clean.csv dataset using R and the tidyverse. Wherever possible, your code should use tidyverse functions.

\--

Complete the following analysis in `R` and generate an `RMarkdown` report to show the analysis and results:

1.  Read in the `gapminder_clean.csv` data as a `tibble` using `read_csv`.

    ```{r, message=F, warning=F}
    library(tibble)
    gap.data <- read.csv('gapminder_clean.csv')

    # transform w/ tibble, minus X index col
    gap.data <- tibble(gap.data[,-1]) 
    ```

2.  Filter the data to include only rows where `Year` is `1962` and then make a scatter plot comparing `'CO2 emissions (metric tons per capita)'` and `gdpPercap` for the filtered data.

    ```{r, message=F, warning=F}
    library(tidyverse)
    library(ggpubr)

    # filter for Year == 1962
    gap.filt <- gap.data %>% filter(Year == 1962)

    # scatter plot of CO2 emissions vs. gdpPercap
    plt <- ggplot(gap.filt, 
           aes(x=CO2.emissions..metric.tons.per.capita., 
               y=gdpPercap)) + 
      geom_point() + 
      xlab('CO2 emissions (metric tons per capita)') + 
      ylab('GDP Per Capita') + 
      ggtitle('CO2 Emissions vs. GDP Per Capita')


    plt <- ggscatter(gap.filt, 
                     x=CO2.emissions..metric.tons.per.capita.,
                     y=gdpPercap, add = "reg.line", conf.int = TRUE, 
                     cor.coef = TRUE, cor.method = "pearson",
                     xlab = 'CO2 emissions (metric tons per capita)', 
                     ylab = 'GDP Per Capita')
    plt
    ```

3.  On the filtered data, calculate the correlation of `'CO2 emissions (metric tons per capita)'` and `gdpPercap`. What is the correlation and associated p value?

    ```{r}
    # calculate correlation
    cor.test(gap.filt$CO2.emissions..metric.tons.per.capita., 
        gap.filt$gdpPercap, method = 'pearson')
    ```

4.  On the unfiltered data, answer "In what year is the correlation between `'CO2 emissions (metric tons per capita)'` and `gdpPercap` the strongest?" Filter the dataset to that year for the next step...

5.  Using `plotly`, create an interactive scatter plot comparing `'CO2 emissions (metric tons per capita)'` and `gdpPercap`, where the point size is determined by `pop` (population) and the color is determined by the `continent`. You can easily convert any `ggplot` plot to a `plotly` plot using the `ggplotly()` command.

Now, without further guidance, use your `R` Data Science skills (and appropriate statistical tests) to answer the following (*use the unfiltered dataset*):

1.  What is the relationship between `continent` and `'Energy use (kg of oil equivalent per capita)'`? (stats test needed)
2.  Is there a significant difference between Europe and Asia with respect to `'Imports of goods and services (% of GDP)'` in the years after 1990? (stats test needed)
3.  What is the country (or countries) that has the highest `'Population density (people per sq. km of land area)'` across all years? (i.e., which country has the highest average ranking in this category across each time point in the dataset?)
4.  What country (or countries) has shown the greatest increase in `'Life expectancy at birth, total (years)'` between 1962 and 2007?
